---
title: "Titanic"
author: "Matteo Morella"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    css: "custom.css"
date: '`r format(Sys.time(), "%d %B, %Y")`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 7, fig.width =  15, fig.align = "center")
```


# Packages and dataset

```{r Packages}
library(gtsummary)
library(dplyr)
library(ggplot2)
library(viridis)
library(hrbrthemes)
library(naniar)
library(tidyr)
library(pROC)
library(caret)
library(MASS)
library(klaR)
library(ranger)
library(xgboost)
library(e1071)
library(nnet)
library(missForest)
```


```{r Data}
data_train <- read.csv("~/Kaggle_Titanic/train.csv")
data_test <- read.csv("~/Kaggle_Titanic/test.csv")
gender_submission <- read.csv("~/Kaggle_Titanic/gender_submission.csv")
head(data_train, 5)
head(data_test, 5)
```
# Data description

**survival**	Survival	0 = No, 1 = Yes

**pclass**	Ticket class	1 = 1st, 2 = 2nd, 3 = 3rd: 1st = Upper; 2nd = Middle; 3rd = Lower

**sex**	Sex	

**Age**	Age in years: Age is fractional if less than 1. If the age is estimated, is it in the form of xx.5

**sibsp**	# of siblings / spouses aboard the Titanic: The dataset defines family relations in this way...
Sibling = brother, sister, stepbrother, stepsister
Spouse = husband, wife (mistresses and fiancés were ignored)

**parch**	# of parents / children aboard the Titanic: The dataset defines family relations in this way...
Parent = mother, father
Child = daughter, son, stepdaughter, stepson
Some children travelled only with a nanny, therefore parch=0 for them.

**ticket**	Ticket number	

**fare**	Passenger fare	

**cabin**	Cabin number	

**embarked**	Port of Embarkation	C = Cherbourg, Q = Queenstown, S = Southampton


```{r Modifiche dati}
data_train <- data_train %>%
  mutate(across(c(Survived, Pclass, Sex, Embarked), as.factor))

data_test <- data_test %>%
  mutate(across(c(Pclass, Sex, Embarked), as.factor))
```


```{r Table summary}

tbl_summary(data_train, include = c(Survived, Pclass, Sex, SibSp, Parch, Embarked))
tbl_summary(data_test, include = c(Pclass, Sex, SibSp, Parch, Embarked))

```

## Age distribution


```{r Histogram age}
ggplot(data_train, aes(x = Age)) +
  geom_histogram(aes(y = ..density..), fill = "#69b3a2", color = "black", alpha = 0.7, bins = 30) +
  geom_density(color = "blue", size = 1.2, linetype = "dashed") + # Linea di densità sovrapposta
  stat_bin(
    aes(y = ..density.., label = ifelse(..density.. == max(..density..), round(..x.., 1), "")),
    geom = "text", vjust = -0.5, bins = 30, color = "black", size = 4
  ) +
  labs(title = "Distribution of Age",
       x = "Age",
       y = "Density") +
  theme_bw(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),  # Centra e grassetta il titolo
    axis.title = element_text(face = "bold"),               # Titoli degli assi in grassetto
    axis.text = element_text(color = "darkblue")            # Colore del testo degli assi
  )

```

```{r Histogram age ~ sex}

ggplot(data_train, aes(x = Age, fill = Sex)) +
  geom_histogram(aes(y = ..density..), position = "identity", alpha = 0.7, bins = 30, color = "black") + 
  geom_density(aes(y = ..density..), color = "blue", size = 1.2, linetype = "dashed", fill = NA) +  # Solo linea tratteggiata
  labs(title = "Distribution of Age by Sex",
       x = "Age",
       y = "Density") +
  scale_fill_manual(values = c("darkorchid2", "darkturquoise")) +  # Colori personalizzati
  facet_wrap(~ Sex, scales = "free_x") +  # Faceting per sesso
  theme_bw(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),  # Centra e grassetta il titolo
    axis.title = element_text(face = "bold"),               # Titoli degli assi in grassetto
    axis.text = element_text(color = "darkblue")            # Colore del testo degli assi
  )

```

```{r age ~ class}
ggplot(data_train, aes(x = Age, fill = Pclass)) +
  geom_histogram(aes(y = ..density..), position = "identity", alpha = 0.7, bins = 30, color = "black") + 
  geom_density(aes(y = ..density..), color = "blue", size = 1.2, linetype = "dashed", fill = NA) +  # Solo linea tratteggiata
  labs(title = "Distribution of Age by socio-economic status",
       x = "Age",
       y = "Density") +
  scale_fill_manual(values = c("darkorchid2", "darkturquoise", "orange")) +  # Colori personalizzati
  facet_wrap(~ Pclass, scales = "free_x") +  # Faceting per sesso
  theme_bw(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),  # Centra e grassetta il titolo
    axis.title = element_text(face = "bold"),               # Titoli degli assi in grassetto
    axis.text = element_text(color = "darkblue")            # Colore del testo degli assi
  )
```

## Fare distribution


```{r hist fare}

ggplot(data_train, aes(x = Fare)) +
  geom_density(aes(y = ..density..), fill = "darkgreen", alpha = 0.7) +
  labs(title = "Density of Fare",
       x = "Fare",
       y = "Density") +
  theme_bw(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),  # Centra e grassetta il titolo
    axis.title = element_text(face = "bold"),               # Titoli degli assi in grassetto
    axis.text = element_text(color = "darkblue")            # Colore del testo degli assi
  )
```

```{r fare ~ class}

ggplot(data_train, aes(x = Fare, fill = Pclass)) +
  geom_histogram(aes(y = ..density..), position = "identity", alpha = 0.7, bins = 30, color = "black") + 
  labs(title = "Distribution of Fare by socio-economic status",
       x = "Fare",
       y = "Density") +
  scale_fill_manual(values = c("darkorchid2", "darkturquoise", "orange")) +  # Colori personalizzati
  facet_wrap(~ Pclass, scales = "free_x") +  # Faceting per sesso
  theme_bw(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),  # Centra e grassetta il titolo
    axis.title = element_text(face = "bold"),               # Titoli degli assi in grassetto
    axis.text = element_text(color = "darkblue")            # Colore del testo degli assi
  )

```

```{r Box-plot fare}
data_train %>%
  ggplot( aes(x = Pclass, y = Fare, fill = Pclass)) +
       geom_boxplot() +
    scale_fill_viridis(discrete = TRUE, alpha = 0.8) +
    geom_jitter(color="black", size = 0.3, alpha = 2) +
    theme_ipsum() +
    theme(
      legend.position="none",
      plot.title = element_text(size=14)
    ) +
    ggtitle("Boxplot for Fare") +
    xlab("Status")
```

# Handling missing values

```{r}
gg_miss_var(data_train)
```

```{r}
gg_miss_var(data_test)
```


```{r}
set.seed(1234)

data_train_impute <- data_train[, c("Survived", "Pclass", "Sex", "Age", "SibSp", "Parch", "Fare"), drop = FALSE]
imputed_data_train <- missForest(data_train_impute)
new_data_train <- imputed_data_train$ximp
dim(new_data_train)
dim(data_train)
table(is.na(new_data_train$Age))


data_test_impute <- data_test[, c("Pclass", "Sex", "Age", "SibSp", "Parch", "Fare"), drop = FALSE]
imputed_data_test <- missForest(data_test_impute)
new_data_test <- imputed_data_test$ximp
table(is.na(new_data_test$Age))
table(is.na(new_data_test$Fare))

```


# Data modeling

## Logistic regression
```{r Logit}

mod_logit1 <- glm(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare, data = new_data_train, family = binomial)

summary(mod_logit1)

predict_logit1 <- predict.glm(mod_logit1, newdata = new_data_train, type = "response")

roc_logit1 <- roc(new_data_train$Survived ~ predict_logit1, plot = T)
legend("bottomright", legend = paste("AUC =", round(auc(roc_logit1), 4)), lwd = 2, box.lwd = 0, bg = "white")


confusione1.1 <- confusionMatrix(as.factor(ifelse(predict_logit1>0.5, 1, 0)), as.factor(new_data_train$Survived)) 
confusione1.1

```

```{r Stepwise}

step1 <- step(mod_logit1, direction = "both", scope = .~(.)^2, trace = 2)

```


```{r Logit interazioni}
mod_logit2 <- glm(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Pclass:Sex + 
    Pclass:Parch + Age:SibSp + Age:Parch + Age:Fare + Pclass:Fare + 
    Sex:Age, data = new_data_train, family = binomial)

summary(mod_logit2)

predict_logit2 <- predict.glm(mod_logit2, newdata = new_data_train, type = "response")

roc_logit2 <- roc(new_data_train$Survived ~ predict_logit2, plot = T)
legend("bottomright", legend = paste("AUC =", round(auc(roc_logit2), 4)), lwd = 2, box.lwd = 0, bg = "white")


confusione1.2 <- confusionMatrix(as.factor(ifelse(predict_logit2>0.5, 1, 0)), as.factor(new_data_train$Survived)) 
confusione1.2
```

## Discriminant analysis

```{r LDA}
mod_lda <- lda(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare, data = new_data_train)

summary(mod_lda)

predict_lda <- predict(mod_lda, newdata = new_data_train)
lda_class <- predict_lda$class
lda_values <- predict_lda$posterior[, 2]

roc_lda <- roc(new_data_train$Survived ~ lda_values, plot = T)
legend("bottomright", legend = paste("AUC =", round(auc(roc_lda), 4)), lwd = 2, box.lwd = 0, bg = "white")

confusione_lda <- confusionMatrix(as.factor(lda_class), as.factor(new_data_train$Survived))
confusione_lda


```

```{r RDA}

mod_rda <- rda(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare, data = new_data_train)

summary(mod_rda)

predict_rda <- predict(mod_rda, newdata = new_data_train)
rda_class <- predict_lda$class
rda_values <- predict_lda$posterior[, 2]

roc_rda <- roc(new_data_train$Survived ~ rda_values, plot = T)
legend("bottomright", legend = paste("AUC =", round(auc(roc_lda), 4)), lwd = 2, box.lwd = 0, bg = "white")

confusione_rda <- confusionMatrix(as.factor(rda_class), as.factor(new_data_train$Survived))
confusione_rda

```

## Tree based Model

```{r Random Forest}
set.seed(1234)
mod_rf <- ranger(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare,
                     data = new_data_train, num.trees = 500, mtry = 3, 
                     seed = 42, probability = T)

mod_rf_predict <- ranger(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare,
                     data = new_data_train, num.trees = 500, mtry = 3, 
                     seed = 42)
summary(mod_rf)


predict_rf1 <- predict(mod_rf, data = new_data_train)
predict_rf2 <- predict_rf1$predictions[, 2]
class_predict_rf <- predict(mod_rf_predict, data = new_data_train)

confusione_rf <- confusionMatrix(as.factor(class_predict_rf$predictions), as.factor(new_data_train$Survived))
confusione_rf

roc_rf <- roc(new_data_train$Survived ~ as.numeric(predict_rf2), plot = T)
legend("bottomright", legend = paste("AUC =", round(auc(roc_rf), 4)), lwd = 2, box.lwd = 0, bg = "white")

       
```

```{r Bagging}

set.seed(5678)
mod_bagging <- ranger(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare,
                     data = new_data_train, num.trees = 500, mtry = 6, seed = 42, probability = T)
mod_bagging_pred <- ranger(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare,
                     data = new_data_train, num.trees = 500, mtry = 6, seed = 42)

predict_bagging <- predict(mod_bagging, data = new_data_train)
predict_bagging.1 <- predict_bagging$predictions[, 2]
predict_bagging_class <- predict(mod_bagging_pred, data = new_data_train)

confusione_bagging <- confusionMatrix(as.factor(predict_bagging_class$predictions), as.factor(new_data_train$Survived))
confusione_bagging

roc_bagging <- roc(new_data_train$Survived ~ as.numeric(predict_bagging.1), plot = T)
legend("bottomright", legend = paste("AUC =", round(auc(roc_bagging), 4)), lwd = 2, box.lwd = 0, bg = "white")

```

## Neural network

```{r Neural Net}
mod_nnet <- nnet(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare, data = new_data_train, size = 10, maxit = 1000, linout = FALSE)

predict_nnet <- predict(mod_nnet, data = new_data_train)

confusione_nnet <- confusionMatrix(as.factor(ifelse(predict_nnet >0.5, 1, 0)), as.factor(new_data_train$Survived))
confusione_nnet

roc_nnet <- roc(new_data_train$Survived ~ as.numeric(predict_nnet), plot = T)
legend("bottomright", legend = paste("AUC =", round(auc(roc_nnet), 4)), lwd = 2, box.lwd = 0, bg = "white")

```

## Model selection

```{r}
accuracy_comparison <- cbind(
  round(confusione1.1$overall["Accuracy"], 5), 
  round(confusione1.2$overall["Accuracy"], 5), 
  round(confusione_lda$overall["Accuracy"], 5), 
  round(confusione_rda$overall["Accuracy"], 5), 
  round(confusione_rf$overall["Accuracy"], 5), 
  round(confusione_bagging$overall["Accuracy"], 5), 
  round(confusione_nnet$overall["Accuracy"], 5)
)

colnames(accuracy_comparison) <- c("Logit", "Logit2", "LDA", "RDA", "RandomForest", "Bagging", "NeuralNetwork")

roc_comparison <- cbind(
  round(auc(roc_logit1), 5), 
  round(auc(roc_logit2), 5), 
  round(auc(roc_lda), 5), 
  round(auc(roc_rda), 5), 
  round(auc(roc_rf), 5), 
  round(auc(roc_bagging), 5), 
  round(auc(roc_nnet), 5))

colnames(roc_comparison) <- c("Logit", "Logit2", "LDA", "RDA", "RandomForest", "Bagging", "NeuralNetwork")

acc_auc <- rbind(accuracy_comparison, roc_comparison)

rownames(acc_auc) <- c("Accuracy", "AUC")
acc_auc
```

# Final result

```{r}

test_resutl_rf <- predict(mod_rf_predict, data = new_data_test)
head(test_resutl_rf$predictions)

test_resutl_bagging <- predict(mod_bagging_pred, data = new_data_test)
head(test_resutl_bagging$predictions)

test_result_nnet <- predict(mod_nnet, newdata = new_data_test, type = "class")
head(test_result_nnet)

```

```{r}

submission_rf <- gender_submission
submission_rf$Survived <- test_resutl_rf$predictions
write.csv(submission_rf, "submission_rf.csv", row.names = F)

submission_bag <- gender_submission
submission_bag$Survived <- test_resutl_bagging$predictions
write.csv(submission_bag, "submission_bagging.csv", row.names = F)

submission_nnet <- gender_submission
submission_nnet$Survived <- test_result_nnet
write.csv(submission_nnet, "submission_nnet.csv", row.names = F)
```

